### 2021-05-01 notes/linear_setup/linear_cartesian_setup.tex
adds writeup of linear conductivity setup

\newcommand\Ra{\mathrm{Ra}}
\newcommand\Pran{\mathrm{Pr}}
\newcommand\Rac{\mathrm{Ra}_{\mathrm{c}}}
\newcommand\Ek{\mathrm{Ek}}
\newcommand\Ro{\mathrm{Ro}}
\newcommand\Nu{\mathrm{Nu}}
\newcommand\Sc{\mathrm{Sc}}

@@

\section{Nondimensional equations}
The nondimensional system of equations we solve is
\begin{align}
&\grad\dot\vec{u} = 0 \\
&\partial_t \vec{u} + \vec{u}\dot\grad\vec{u} = -\grad \varpi + T \hat{z} + \rm{Re}_0^{-1}\grad^2 \vec{u} \\
&\partial_t T + \vec{u}\dot\grad T + w \grad_{\rm{ad}} = \rm{Pe}_0^{-1}\grad^2 T' + \grad\dot[k \grad \overline{T}] + Q
\end{align}
These are just the Boussinesq equations of motion, with some slight tweaks:
\begin{enumerate}
\item There is a nonzero adiabatic temperature gradient, $\partial_z T_{\rm{ad}} = -\grad_{\rm{ad}}$.
We choose this sign convention ($\grad_{\rm{ad}}$ and similarly $\grad_{\rm{rad}}$ will be positive) to align better with the stellar structure community's intuition.
$|\grad T_0|$ and $\grad_{\rm{ad}}$ may be much larger in magnitude than the convective fluctuations.
The background temperature field $T_0$ does not enter into the momentum equation, as it is automatically canceled by a background $\grad \varpi$.
\item Convection is driven by internal heating, $Q$.
I've confirmed with some simple tests that $|u|^2 \approx Q$, and therefore the convective frequency is $f_{\rm{conv}}^2 \approx Q$.
\item The control freefall P\'{e}clet number, $\rm{Pe}_0$, and the diffusivity associated with it, only acts on fluctuations, not on the $k_x = 0$ mode.
For the $k_x = 0$ mode, we model the radiative flux as $F_{\rm{rad}} = -k \grad \overline{T}$, where $k$ is a nonconstant coefficient, $k(z)$.
$k$ may be a larger or smaller diffusivity than $\rm{Pe}_0^{-1}$, and its magnitude is set by $\int Q\, dz \approx -k \grad\overline{T}$.
\end{enumerate}

\section{Model Setup}
We will study a two-layer model in $z = [0, L_z]$ in which the conductivity, $k(z)$ is a step function such that
\begin{equation}
k(z) = \begin{cases}
k_b & z < L_{\rm{cz}} \\
k_t & z \geq L_{\rm{cz}}
\end{cases}
\end{equation}
where subscript ``t'' refers to the top layer and ``b'' refers to the bottom.
Generally, we will set $L_{\rm{cz}} = 1$.
(These won't be true discontinuities, but will be smooth step-like functions achieved with erfs.)

There will be two important input parameters in this system:
\begin{enumerate}
\item The stiffness, 
\begin{equation}
\mathcal{S} = \frac{N_{t}^2}{f_{\mathrm{conv}}^2},
\end{equation}
which is the ratio of the buoyancy frequency (in the stable layer) to the square convective frequency.
\item $\mathcal{P}$, a parameter which sets the penetration depth.
$\mathcal{P}$ determines what the magnitude of the \emph{negative} convective flux would be in an adiabatic penetrative layer with $k = k_t$.
We will define $\mathcal{P} = -F_{\rm{conv}}/F_{\rm{conv, rz}}$, and our expectation is that we get a lot of convective penetration when $\mathcal{P} \gg 1$ and none when $\mathcal{P} \ll 1$.
\end{enumerate}
In addition to these parameters, the magnitude of the heating, $Q$ and the vertical extent of the heating layer, $\delta_H$, must be specified as they set $F_{\rm{conv}} = Q \delta_H$.
We will choose $Q = 1$ and $\delta_H = 0.2$ so that $F_{\rm{conv}} = 0.2$; we will furthermore choose and $F_{\rm{bot}} = \zeta F_{\rm{conv}}$, where $\zeta$ is some arbitrary scalar that sets how much smaller $k$ is in the cz than the RZ; we'll set $\zeta = 10^{-3}$ for now.

Under these choices, the system is determined by five equations,
\begin{enumerate}
\item At some point, the adiabatic gradient is equal to the radiative gradient,
\begin{equation}
k_{\rm{ad}}\grad_{\rm{ad}} = F_{\rm{tot}}.
\end{equation}
\item at $z = 0$, the base of the CZ, the adiabatic temperature gradient is the radiative gradient, $\grad_{\rm{ad}} = \grad_{\rm{rad}}$, and carries $F_{\rm{bot}}$,
\begin{equation}
k_b \grad_{\rm{ad}} = F_{\rm{bot}}
\end{equation}
\item The radiative gradient can carry the flux in the RZ,
\begin{equation}
k_{t} \grad_{\rm{rad,t}} = F_{\rm{tot}}
\end{equation}
\item Assuming $\grad T = -\grad_{\rm{ad}}$ everywhere, the convective flux in the top layer is $F_{\rm{conv,t}} = F_{\rm{conv}} - \Delta k \grad_{\rm{ad}}$, where $F_{\rm{conv}} = Q \delta_H$ and $\Delta k = k_t - k_b$.
Thus, definitionally,
\begin{equation}
\mathcal{P} = -\frac{F_{\rm{conv}}}{F_{\rm{conv,t}}} = \left(\frac{\Delta k \grad_{\rm{ad}}}{Q \delta_H} - 1\right)^{-1}.
\end{equation}
\item Defining $f_{\rm{conv}}^2 \approx Q$ and $N^2 \approx -(\grad_{\rm{rad,t}} - \grad_{\rm{ad}})$, the stiffness gives
\begin{equation}
\mathcal{S} = \frac{N^2}{f_{\mathrm{conv}}^2} = \frac{-(\grad_{\rm{rad,t}} - \grad_{\rm{ad}})}{Q}
= \frac{F_{\mathrm{tot}}}{Q}(k_{\rm{ad}}^{-1} - k_t^{-1})
\end{equation}
\end{enumerate}

Solving this system of equations gives
\begin{align}
k_t = \frac{\delta_H}{\mathcal{S} \mathcal{P}},\qquad
k_b = k_t \frac{\zeta}{1 + \zeta + \mathcal{P}^{-1}},\qquad
k_{\rm{ad}} = k_t \frac{1 + \zeta}{1 + \zeta + \mathcal{P}^{-1}},\\
\grad_{\rm{ad}} = Q \mathcal{S}\mathcal{P}(1 + \zeta + \mathcal{P}^{-1}),\qquad
\grad_{\rm{rad}} = Q \mathcal{S}\mathcal{P}(1 + \zeta).
\end{align}

>>>>

\newcommand{\gradrad}{\ensuremath{\nabla_{\rm{rad}}}}
\newcommand{\gradad}{\ensuremath{\nabla_{\rm{ad}}}}
\newcommand{\Ftot}{\ensuremath{F_{\rm{tot}}}}
\newcommand{\Frad}{\ensuremath{F_{\rm{rad}}}}
\newcommand{\Fconv}{\ensuremath{F_{\rm{conv}}}}
\newcommand{\mP}{\ensuremath{\mathcal{P}}}
\newcommand{\mR}{\ensuremath{\mathcal{R}}}
\newcommand{\mS}{\ensuremath{\mathcal{S}}}
\newcommand\Pran{\ensuremath{\mathrm{Pr}}}
\newcommand\Ra{\mathrm{Ra}}
\newcommand\Rac{\mathrm{Ra}_{\mathrm{c}}}
\newcommand\Ek{\mathrm{Ek}}
\newcommand\Ro{\mathrm{Ro}}
\newcommand\Nu{\mathrm{Nu}}
\newcommand\Sc{\mathrm{Sc}}

@@

\section{Nondimensional equations}
The nondimensional system of equations we solve is
\begin{align}
&\grad\dot\vec{u} = 0 \\
&\partial_t \vec{u} + \vec{u}\dot\grad\vec{u} = -\grad \varpi + T \hat{z} + \mR^{-1}\grad^2 \vec{u} \\
&\partial_t T + \vec{u}\dot\grad T + w \grad_{\rm{ad}} = (\Pran\mR)^{-1}\grad^2 T' + \grad\dot[k \grad \overline{T}] + Q
\end{align}
These are just the Boussinesq equations of motion, with some slight tweaks:
\begin{enumerate}
\item There is a nonzero adiabatic temperature gradient, $\partial_z T_{\rm{ad}} = -\grad_{\rm{ad}}$.
We choose this sign convention ($\grad_{\rm{ad}}$ and similarly $\grad_{\rm{rad}}$ will be positive) to align better with the stellar structure community's intuition.
$|\grad T_0|$ and $\grad_{\rm{ad}}$ may be much larger in magnitude than the convective fluctuations.
The background temperature field $T_0$ does not enter into the momentum equation, as it is automatically canceled by a background $\grad \varpi$.
\item Convection is driven by internal heating, $Q$.
I've confirmed with some simple tests that $|u|^2 \approx Q$, and therefore the convective frequency is $f_{\rm{conv}}^2 \approx Q$.
$|u|$ is also a weak function of $\mR$, but is a very strong function of $Q$.
\item The control freefall Reynolds number (\mR) and P\'{e}clet number (\Pran\mR)---and the diffusivity associated with them---only act on fluctuations, not on the $k_x = 0$ mode.
For the $k_x = 0$ mode, we model the radiative flux as $F_{\rm{rad}} = -k \grad \overline{T}$, where $k$ is a nonconstant coefficient, $k(z)$.
$k$ may be a larger or smaller diffusivity than $\rm{Pe}_0^{-1}$, and its magnitude is set by $\int Q\, dz \approx -k \grad\overline{T}$.
\end{enumerate}

\section{Model Setup}
We will study a two-layer model in $z = [0, L_z]$ in which the gradient of the conductivity, $\partial_z k(z)$, is a step function,
\begin{equation}
\frac{\partial k(z)}{\partial z} = \partial_z k_{0}\begin{cases}
1 & z < L_{\rm{cz}} \\
\mP^{-1} & z \geq L_{\rm{cz}}
\end{cases}
\end{equation}
Generally, we will set $L_{\rm{cz}} = 1$.
The crucial thing here is that the CZ has a nondimensional slope of order $\partial_z k_0$, but above the place where the Schwarzschild criterion is met ($\gradrad = \gradad$) the slope is larger or smaller than that value according to the magnitude of \mP.
$\mP$ therefore sets how quickly the magnitude of $\Frad$ conducted along a (constant) $\gradad$ varies with height.

Aside from $\mR$ and $\Pran$, there will be two important input parameters in this system:
\begin{enumerate}
\item The stiffness, 
\begin{equation}
\mathcal{S} = \frac{N_{\rm{RZ}}^2}{f_{\mathrm{conv}}^2},
\end{equation}
which is the characteristic ratio of the buoyancy frequency (in the stable layer) to the square convective frequency.
Since $N_{\rm{RZ}}^2$ (and $\gradrad$) will vary with height, we will nondimensionalize using its characteristic value at $z = 2 L_{\rm{cz}}$.
\item $\mathcal{P}$, a parameter which sets the penetration depth.
$\mathcal{P}$ determines how steeply the value of $\Frad$ (conducted along $\gradad$ in the penetration zone) changes as a function of height.
As a result, assuming flux conservation with $\Ftot = \Frad + \Fconv$, $\mP$ sets how rapidly the convective flux decreasing (into negative values) as a function of height.
We will define $\mP$ using an atmosphere where $\grad = \gradad$ everywhere; In this atmosphere,
\begin{equation}
\mP \equiv -\frac{\Fconv|_{\rm{cz}}}{\Fconv|_{2L_{\rm{cz}}}}
= -\frac{(\Ftot - \Frad)|_{\rm{cz}}}{(\Ftot - \Frad)|_{2L_{\rm{cz}}}}
\end{equation}
Our expectation is that we get a lot of convective penetration when $\mathcal{P} \gg 1$ and none when $\mathcal{P} \ll 1$.
\end{enumerate}
In addition to these parameters, the magnitude of the heating, $Q$ and the vertical extent of the heating layer, $\delta_H$, must be specified as they set $F_{\rm{conv}} = Q \delta_H$.
We will choose $Q = 1$ and $\delta_H = 0.2$ so that $F_{\rm{conv}} = 0.2$; we will furthermore choose and $F_{\rm{bot}} = \zeta F_{\rm{conv}}$, where $\zeta$ is some arbitrary scalar that sets how much smaller $k$ is in the cz than the RZ; we'll set $\zeta = 10^{-3}$ for now.
This choice sets a boundary condition on the integral of $k_0$ so that $k_0(z=0) = F_{\rm{bot}}/\gradad$.

Under these choices, the system is determined by five equations,
\begin{enumerate}
\item At $z = L_{\rm{cz}}$, the adiabatic gradient is equal to the radiative gradient,
\begin{equation}
k_{\rm{ad}}\grad_{\rm{ad}} = F_{\rm{tot}}.
\end{equation}
\item at $z = 0$, the base of the CZ, the adiabatic temperature gradient is the radiative gradient, $\grad_{\rm{ad}} = \grad_{\rm{rad}}$, and carries $F_{\rm{bot}}$,
\begin{equation}
k_0(z=0) \grad_{\rm{ad}} = F_{\rm{bot}}
\end{equation}
\item The radiative gradient can carry the flux in the RZ,
\begin{equation}
k(z > L_{\rm{cz}}) \grad_{\rm{rad}} = (k_{\rm{ad}} + \partial_z k_0 \mP^{-1}(z - L_{\rm{cz}}))\gradrad = F_{\rm{tot}}
\end{equation}
\item The magnitude of $\mP$ is based on the radiative flux conducted along the adiabat at $z = 2 L_{\rm{cz}}$, where $k(z=2L_{\rm{cz}}) = k_{\rm{ad}} + \partial_z k_0 L_{\rm{cz}}\mP^{-1}$.
Thus, with $\Fconv|_{\rm{cz}} = Q\delta_H$,
\begin{equation}
\mP = -\frac{\Fconv|_{\rm{cz}}}{\Fconv|_{2L_{\rm{cz}}}}
= -\frac{Q\delta_H}{\partial_z k_0 L_{\rm{cz}}\mP^{-1}\gradad}
\Rightarrow
\mP^2 = \frac{Q \delta_H}{\gradad \partial_z k_0 L_{\rm{cz}}}
\end{equation}
\item Defining $f_{\rm{conv}}^2 \approx Q$ and $N^2 \approx -(\grad_{\rm{rad}}|_{2L_{\rm{cz}}} - \grad_{\rm{ad}})$, the stiffness gives
\begin{equation}
\mS = \frac{N^2}{f_{\mathrm{conv}}^2} = \frac{-(\grad_{\rm{rad}}|_{2L_{\rm{cz}}} - \grad_{\rm{ad}})}{Q}
= \frac{F_{\mathrm{tot}}}{Q}(k_{\rm{ad}}^{-1} - [k_{\rm{ad}} + \partial_z k_0 \mP^{-1} L_{\rm{cz}}]^{-1})
\end{equation}
\end{enumerate}

Solving this system of equations gives
\begin{align}
\partial_z k_0 = \frac{\delta_H}{L \mP^2 \mS}\frac{1}{\xi},\qquad
k_b = \frac{\delta }{\mS}\frac{\zeta}{\xi},\qquad
k_{\rm{ad}} = \frac{\delta}{\mS}\frac{1 + \zeta}{\xi},\qquad
\gradad = Q \mS \xi 
\end{align}
with $\xi \equiv 1 + \mP^3(1 + \zeta)$ and $\gradrad = F_{\rm{tot}}/k(z)$.

### 2021-05-01 notes/penetration_argument/penetration_argument.tex
adds writeup of linear conductivity setup

\documentclass[12pt,preprint]{article}
\usepackage{natbib}
\bibliographystyle{plain}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}

@@

@@

>>>>

\documentclass[12pt]{article}
\usepackage{natbib}
\bibliographystyle{aasjournal}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}

@@

\newcommand{\aap}{A\&A}

@@

\section{Zahn argument}
Here we will work through a Boussinesq version of the argument in section 3.1 of \citet{zahn1991}.
In the adiabatic penetrative region, we know that the radiative flux is
\begin{equation}
F_{\rm{rad}} = -k \frac{\partial T_{\rm{ad}}}{\partial z}
\end{equation}
Further, we know that the convective flux is the difference between the total total and radiative flux,
\begin{equation}
F_{\rm{conv}} = F_{\rm{tot}} - F_{\rm{rad}}.
\end{equation}
Furthermore, the convective flux is
\begin{equation}
F_{\rm{conv}} = \overline{w T} = f \, W(z) \, \delta T(z),
\end{equation}
where $f$ is the filling factor of the flows that carry the flux in the penetration zone (upflows) and $W$ and $T$ give the vertical shape of typical velocity and temperature perturbations.

We next assume that the temperature and velocity are highly correlated in the overshooting convective motions (upflows).
This allows us to describe the horizontal nature of all fields in upflows using a function $h(x,y)$ so that
\begin{equation}
w(x,y,z) = W(z) h(x,y)
\qquad
T'(x,y,z) = \delta T(z) h(x,y),
\end{equation}
where $T'$ are the temperature fluctuations away from the mean.
By continuity and by our above definition of the convective flux, we have
\begin{equation}
\overline{h} = 0,
\qquad
\overline{h^2} = f.
\end{equation}
